---
title: "Week4.BivariateDist.Categ.Data"
author: "Anita Owens"
date: "March 10, 2016"
output: html_document
---
Week 4: Bivariate Distributions (Categorical Data)

For artists age 30 or older, do female artists play different kinds of music on Austin City Limits than male artists?
```{r}
setwd("/Users/anitaowens/Documents/R")
```
read csv
```{r}
acl<-read.csv("AustinCityLimits.csv")
names(acl)
```
Create a subset of the data for artists age 30 or older.
```{r}
over30<- acl[ which(acl$Age >=30), ]

older <-acl[acl$Age>=30,]
```
Create a table to show the marginal distribution for each variable. 
```{r}
genre<-table(over30$Genre)
genre

gender<-table(over30$Gender)
gender
```
# Create contingency table 
```{r}
twoway <- table(over30$Gender,over30$Genre)
twoway
addmargins(twoway)
```
# Visualize the counts
```{r}
#side by side
barplot(twoway, legend=T, beside=T, col = c( "pink", "blue"), main = "Austin City Limits Music Genre by Gender")

#stacked
barplot(twoway, legend=TRUE, col = c( "pink", "blue"), main = "Austin City Limits Music Genre by Gender")

```
# Calculate P(A): the probability of each genre being played
```{r}
prop.table(genre)
```
# Calculate P(A|B): the probability of each genre being played, given the artist’s gender
```{r}
prop.table(twoway,1)
prop.table(twoway)
```
How many male artists are in the 30+ year old artist subset of the Austin City Limits dataset?
```{r}
install.packages('plyr')
library(plyr)
count(over30, "Gender")

#or
gender
```
To determine the proportion of jazz performers that were male, you would ...
```{r}
twoway
#divide 7 by 11
round(7/11, digits = 3)
```
To determine the proportion of males that performed jazz, you would divide 
```{r}
round(7/77, digits = 3)
```
Which of these lines of code provides the probability that a randomly selected artist from the dataset performed rock/folk/indie music?
```{r}
prop.table(genre)
```
What is the probability that a randomly selected artist from the dataset performed rock/folk/indie music?
```{r}
round(61/102, digits = 3)

#you can also access the probability with the genre table
prop.table(genre)
```
Which of these lines of code provides the probability that a randomly selected female artist performed rock/folk/indie music?
```{r}
#the 1 tells R to compare the sexes in the rows.
prop.table(twoway, 1)
```
What is the probability that a randomly selected female artist performed rock/folk/indie music?
```{r}
?round(25/61, digits = 4)
#.320
```
Lab Questions
1. Subset the data (males only).
```{r}
males<-acl[ which(acl$Gender=="M"), ]
```
2. Create a table to show the marginal distributions for Genre and Grammy.
```{r}
grammy<-table(males$Grammy)
grammy

genre<-table(males$Genre)
genre
```
3. Create a contingency table to show the conditional distribution for Genre and Grammy.
```{r}
gg<-table(males$Grammy, males$Genre)
gg
```
4. Make a bar chart to better visualize how many artists in each Genre received a Grammy.
```{r}
barplot(gg, legend=T, beside=T, main = "How many artists in each genre won grammys", col = c("red", "blue"))
```
5. Calculate P(A):  the probability of winning a Grammy.
```{r}
prop.table(grammy)

#43%
```
6. Calculate P(A|B): the probability of winning Grammy, given the artist's Genre.
```{r}
prop.table(gg)
```
7. Interpret what these probabilities tell us about the relationship between Genre and winning a Grammy.
```{r}

```
1) How many male artists won a Grammy?
2) How many male artists did not win a Grammy?
```{r}
grammy
#35
#46
```
3) Which genre had the greatest number of Grammy wins?
```{r}
genre
#Rock/Folk/Indie 56
```
4) What is the probability that a randomly selected artist was a Grammy winner? (Report as a proportion rounded to three decimal places)
```{r}
prop.table(grammy)
#or
round(35/81, digits = 3)
```
5) To determine the probability of winning a Grammy if the artist was a singer-songwriter, you would divide _______ by ________. (Enter numerical values.)
```{r}
addmargins(grammy)
addmargins(gg)

round(2/7, digits = 3)
#[1] [1] 0.286
```
6) To determine the probability that a randomly-selected Grammy winner was a singer-songwriter, you would divide ________ by ________. (Enter numerical values.)
```{r}
prop.table(gg, 1)
#answer:  0.05714286
2/35

#num of grammy winners in songwriting category divided
#by the total number of grammy winners
round(2/35, digits = 3)

```
Conditional Probabilities

What is the probability that a randomly selected male artist from each of the following genres won a Grammy? (Report as proportions rounded to three decimal places.)

```{r}
#Country
round(7/11, digits = 3)

#jazz
round(4/7, digits = 3)

#rock
round(22/56, digits = 3)

#singer/songwriter
round(2/7, digits = 3)

#7e. Our conclusion should be that winning a Grammy _______ independent of Genre.
#Answer: is not
```
Problem Set
#1 1.  Generate a table to show the number of artists that are "popular" and those that are not.
```{r}
pop<-table(acl$Facebook.100k)
pop

1=pop
0=not popular
```
Generate a table to show the number of "popular" artists within each age group.
```{r}
age<-table(acl$Facebook.100k, acl$Age.Group)
age
```

