---
title: "data first"
author: "Anita Owens"
date: "April 8, 2016"
output: ioslides_presentation
---

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Code and Output

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```

read file
```{r}
employees <- read.csv("http://www.headfirstlabs.com/books/hfda/hfda_ch09_employees.csv",header=TRUE)
```

## I want to look at this data set's structure

```{r}
str(employees)
```
##Histogram of employees who received a raise
```{r}
hist(employees$received, main = "Frequency of Percentage of Raises")
```
##with groupings
```{r}
hist(employees$received, breaks=50, main = "Frequency of Percentage of Raises")
```
##Explore further
```{r}
dim(employees)
sd(employees$received)
summary(employees$received)
summary.data.frame(employees)
```
##2007
```{r}
hist(employees$received[employees$year==2007], breaks=50)
hist(employees$received[employees$gender=="F"], breaks=50, main = "Female Employees")
```

