---
title: "dplyr"
author: "Anita Owens"
date: "December 25, 2015"
output: html_document
---
```{r}
install.packages("nycflights13")
library(nycflights13)
dim(flights)
head(flights)

```
Filter rows
```{r}
library(dplyr)
#select all flights on Jan 1st
filter(flights, month == 1, day == 1)
#equivalent to
flights[flights$month == 1 & flights$day == 1, ]

#put subsetted data into a new data frame
june<-filter(flights, month==6, day==20)

```
To select rows by position, use slice()
```{r}
slice(flights, 1:10)
```
Select colums with select
```{r}
# Select columns by name
select(flights, year, month, day)

#put selected column into a new dataframe
orig<-select(flights, origin)
```
Filter rows with filter()
Equivalent to: flights[flights$month == 1 & flights$day == 1, ]
```{r}
filter(flights, month == 1, day == 1)
```
Single table verbs

Dplyr aims to provide a function for each basic verb of data manipulation:

filter() (and slice())
arrange()
select() (and rename())
distinct()
mutate() (and transmute())
summarise()
sample_n() and sample_frac()
```{r}
```
To select rows by position, use slice():
```{r}
slice(flights, 1:10)
```
Extract distinct (unique) rows

A common use of select() is to find the values of a set of variables. This is particularly useful in conjunction with the distinct() verb which only returns the unique values in a table.
```{r}
distinct(select(flights, tailnum))
distinct(select(flights, origin, dest))
```
Summarise values with summarise()

The last verb is summarise(). It collapses a data frame to a single row (this is exactly equivalent to plyr::summarise()):
```{r}
summarise(flights,delay = mean(dep_delay, na.rm= TRUE))
```


