---
title: "Import Data Using Packages"
author: "Anita Owens"
date: "August 27, 2016"
output: html_document
---
```{r}
#install once
install.packages("devtools")

#install weather data from github
devtools::install_github("ozagordi/weatherData")
devtools::install_github("repositoryName/packageName")```
```{r}
```
install from CRAN option
```{r}
install.packages('weatherData')
library(weatherData)
```
sample weather data
```{r}
weatherdata <- getWeatherForDate("BOS", "2016-08-01", end_date="2016-08-15")
```
Quickstart Example 1. Start by getting the temperature data for a Location of interest for a particular date.
```{r}
 getWeatherForDate("SEA", "2014-05-05")
```
 If you want detailed data for a range of dates instead of for a single date, just specify a start date and end date:
 
```{r}
 getWeatherForDate("SEA", "2014-04-01", end_date="2014-04-30")
 #This will fetch a dataframe of temperature data for the month of April.
```
Just get me everything for a whole year for a Location of interest.
```{r}
 dfw_wx <- getWeatherForYear("DFW", 2013)
```
Find the station id for the location(s) that you are interested in. If you know the airport code (3 letters) you can try that. Once you have the station_id and the date ranges, it is a simple matter to fetch the data.
```{r}
 jax_wx <- getWeatherForYear("JAX", 2016)
```
First, we get two data frames, with one year's worth of temperature data in each. (365 rows each.) Note that getWeatherForYear is a function in the weatherData package.
```{r}
city1 <- "ORD"
city2 <- "SFO"
df1 <- getWeatherForYear(city1, 2013)
df2 <- getWeatherForYear(city2, 2013)
```
Given two dataframes that contain daily temperatures (mean, min and max), we are simply taking the vectorized differences and creating another data frame.
```{r}
getDailyDifferences <- function(df1, df2){
  Delta_Means <- df1$Mean_TemperatureF - df2$Mean_TemperatureF
  Delta_Max <- df1$Max_TemperatureF - df2$Max_TemperatureF
  Delta_Min <- df1$Min_TemperatureF - df2$Min_TemperatureF

  diff_df <- data.frame(Date=df1$Date, Delta_Means, Delta_Max, Delta_Min)
  return(diff_df)
}
```
Next, let's write a small function using ggplot to plot the daily differences. We will use 3 facets, one each for Mean, Max and Min difference.
```{r}
plotDifferences <- function (differences, city1, city2) {
  library(reshape2)
  m.diff <- melt(differences, id.vars=c("Date"))
  p <- ggplot(m.diff, aes(x=Date, y=value)) + geom_point(aes(color=variable)) +  
    facet_grid(variable ~ .) +geom_hline(yintercept=0)
  p <- p + labs(title=paste0("Daily Temperature Differences: ", city1, " minus ",city2))
  print(p)
}
```
plot
```{r}
differences<- getDailyDifferences(df1, df2)
plotDifferences(differences, city1, city2)
```

