---
title: "inferential.stats"
author: "Anita Owens"
date: "May 8, 2016"
output: html_document
---
```{r Load Libraries}
library(data.table)

```



```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```



import data
```{r}
getwd()
Kloutscores1 <- read.table("~/Documents/GitHub/data/Kloutscores1.csv", quote="\"", comment.char="")

dim(Kloutscores1)
```
mean and standard deviation
```{r}
mean(Kloutscores1$V1)
var(Kloutscores1$V1)
sd(Kloutscores1$V1)
standarddev<-sd(Kloutscores1$V1)
standarddev
```
plot histogram
```{r}
hist(Kloutscores1$V1, col="lightblue")
abline(v=mean(Kloutscores1$V1), col="red", lwd=3, lty=2)
```
histogram with density plot
```{r}
hist(Kloutscores1$V1, col="lightblue")
abline(v=mean(Kloutscores1$V1), col="red", lwd=3, lty=2)
lines(density(Kloutscores1$V1, na.rm = TRUE), col="blue")

#turn data frame column values into a vector for line density plot
avector<-as.vector(Kloutscores1[,1])
class(avector)

library(ggplot2)

# Histogram overlaid with kernel density curve
meantitle<-paste("Mean of Sample Distribution =", round(mean(avector), digits = 3))

ggplot(Kloutscores1, aes(x=avector)) +
        geom_histogram(aes(y=..density..),
                       binwidth = 5,
                       colour="black", fill="white") +
        geom_density(alpha=.2, fill="#FF6666") + # Overlay with transparent density plot
        geom_vline(aes(xintercept=mean(avector, na.rm=T)),# Ignore NA values for mean
               color="red", linetype="dashed", size=1) +
        ggtitle(paste0(meantitle))
        
```

sample mean
http://msemac.redwoods.edu/~darnold/math15/spring2013/R/Activities/DistributionOfSampleMeans.html
http://www.r-bloggers.com/normal-distribution-functions/
```{r}
x<-rnorm(35, 37.71905, 16.04431)
x
mean(x)
hist(x,probability = TRUE)
sd(x)

#What is the standard deviation of the sample population distribution?
#answer
sqrt(35)
16.04431/5.91608
sdpopmean<-sd(Kloutscores1$V1)/sqrt(35)
sdpopmean
```
What's the zscore for a mean of 40?
z= x−μ/σ 
z=score-mean(pronounced mu)/ standard deviation (pronounced sigma)

Sample size of 35. Distribution of sample means (Sample Distribution) m=37.72 

σ/n=2.71

```{r}
zscore<-(40-37.72)/2.71
#or
zscore<-(40-mean(Kloutscores1$V1))/sdpopmean
round(zscore, digits = 3)
```
https://onlinecourses.science.psu.edu/stat200/node/148

Probability of drawing a sample with the mean of 40?

distsamplemean<-standarddev/sqrt(35)
z<-(40-mean(Kloutscores1$V1))/distsamplemean
```{r}
distsamplemean<-standarddev/sqrt(35)
z<-(40-mean(Kloutscores1$V1))/distsamplemean
z
#zscore: 0.8410634
#or
z<-(40-37.71905)/(16.04431/sqrt(35))
z
ztable-.7996
#answer: 1-.7996
[1] 0.2004
d<-density(Kloutscores1$V1)
plot(d)
abline(v=40, col="red")
```
increasing the sample size

1st find the standard error
SE=σ /sqrt(n)
```{r}
se<-standarddev/sqrt(250)
se
distsamplemean<-standarddev/sqrt(250)
distsamplemean
zscore<-(40-mean(Kloutscores1$V1))/distsamplemean
zscore
#[1] 2.247831
ztable probability: 0.9875

```
Probability of randomly selecting a sample of 250 with a mean of at least 40?
```{r}
1-0.9875
```

