---
title: "Lattice Plotting System"
author: "Anita Owens"
date: "January 14, 2016"
output: html_document
---
#set directory
```{r}
setwd("/Users/anitaowens/Documents/datasciencecoursera")
getwd()
```
#load library
```{r}
library(lattice)
```
Lattice Functions

xyplot: main function for creating scatterplots
bwplot: box-and-whiskers plots ("boxplots")
histogram: histograms
stripplot: like a boxplot but with actual points
dotplot: plot dots on "violin strings"
splom: scatterplot matrix, like pairs in base plotting system
levelplot, countourplot: for plotting "image" data

#lattice formula
xyplot(y ~ x | f * g, data)

~ is the formula notation

on the left of the ~ is the y-axis variable, on the right of the ~ is the x-axis variable

f and g are conditioning variables--they are optional
-the * indicates an interaction between 2 variables

the second argument is the data frame or list from which the variables in the formula should be looked up

if no data frame or list is passed, then the parent frame is used

if no other arguments are passed, there are defaults that can be used.

```{r}
#simple scatterplot
library(datasets)
xyplot(Ozone ~ Wind, data=airquality)
```
Lattice Behavior
```{r}
#load library
library(lattice)

#nothing happens
p<-xyplot(Ozone ~ Wind, data=airquality)

#plot appears
print (p)

#autoprinting feature
xyplot(Ozone ~ Wind, data=airquality)

library(datasets)
```
Lattice Panel Functions
-lattice functions have a panel function which controls what happens inside each panel plot
-the lattice package comes with default panel functions, but you can supply your own to customize the panel
-panel functions receive the x/y coordinates of the data points in their panel (along with any optional arguments)
```{r}
set.seed(10)
x<-rnorm(100)
f<-rep(0:1, each=50)
y<-x + f - f * x + rnorm(100, sd=.5)
f<-factor(f, labels = c("Group 1", "Group 2"))

#plot with 2 panels
xyplot(y ~ x | f, layout = c(2,1))
```
Custom panel function
```{r}
xyplot (y ~ x | f, panel = function(x,y,...){
        panel.xyplot(x,y,...)##first call the default panel function for xyplot
        panel.abline(h=median(y), lty=2) ##add a horizontal line at the median
})
```
Lattice Panel Functions: Regression Line
#you can't use base plotting functions in lattice
```{r}
xyplot (y ~ x | f, panel = function(x,y,...){
        panel.xyplot(x,y,...)##first call the default panel function for xyplot
        panel.lmline(x,y, col=2) ##overlay a simple regression line
})
```
Summary

-the lattice system lets you look at a lot of data at once
-lattice plots are constructed with a single function call to a core lattice function (e.g. xyplot)
-aspects like margins and spacing are automatically handled and defaults are usually sufficient
-the lattice system is ideal for creating conditioning plots where you examine the same kind of plot under many different conditions
-panel functions can be specified/customized to modify what is plotted in each of the plot panels

SWIRL

Now try running xyplot with the formula Ozone~Wind as the first argument and the second argument data set equal to airquality.
```{r}
xyplot(Ozone ~ Wind, data=airquality)
```
 Now rerun xyplot with the formula Ozone~Wind as the first argument and the second argument data set equal to
airquality (use the up arrow to save typing). This time add the arguments col set equal to "red", pch set equal to 8, and main set equal to "Big Apple Data".
```{r}
xyplot(Ozone ~ Wind, data=airquality, col="red", pch=8, main= "Big Apple Data")
```
 Run xyplot with the formula Ozone~Wind | as.factor(Month) as the first argument and the second argument data set equal to airquality (use the up arrow to save typing). So far, not much is different, right? Add a third argument, layout,
| set equal to c(5,1).

```{r}
xyplot(Ozone ~ Wind | as.factor(Month), data=airquality, layout= c(5,1))
xyplot(Ozone ~ Wind | Month, data=airquality, layout= c(5,1))
```

```{r}
p<-xyplot(Ozone~Wind,data=airquality)
print(p)
names(p)
mynames[myfull]
p[["formula"]]
p[["x.limits"]]
```

```{r}
xyplot(y~x|f, layout=c(2,1))

#the call to pathfile
source(pathtofile("plot1.R"), local=TRUE)

myedit("plot2.R")
source(pathtofile("plot2.R"), local=TRUE)
```

```{r}
table(diamonds$color, diamonds$cut)
myedit("myLabels.R")
source(pathtofile("myLabels.R"), local=TRUE)
xyplot(price~carat | color*cut, data=diamonds, strip = FALSE, pch=20, xlab = myxlab, ylab = myylab, main=mymain)

#without the strip argument
xyplot(price~carat | color*cut, data=diamonds, pch=20, xlab = myxlab, ylab = myylab, main=mymain)

```
http://www.statmethods.net/graphs/scatterplot.html
```{r}
 # Scatterplot Matrices from the glus Package 
library(gclus)
dta <- mtcars[c(1,3,5,6)] # get data 
dta.r <- abs(cor(dta)) # get correlations
dta.col <- dmat.color(dta.r) # get colors
# reorder variables so those with highest correlation
# are closest to the diagonal
dta.o <- order.single(dta.r) 
cpairs(dta, dta.o, panel.colors=dta.col, gap=.5,
main="Variables Ordered and Colored by Correlation" )

```
High Density Scatterplots
```{r}
# High Density Scatterplot with Binning
library(hexbin)
x <- rnorm(1000)
y <- rnorm(1000)
bin<-hexbin(x, y, xbins=50) 
plot(bin, main="Hexagonal Binning")
```
3D Scatterplots
```{r}
# Spinning 3d Scatterplot
library(rgl)

plot3d(wt, disp, mpg, col="red", size=3)

dev.copy(png, file="3dscatter.png",width=480,height=480)
dev.off()
```
# Multiple Linear Regression Example 
fit <- lm(y ~ x1 + x2 + x3, data=mydata)
summary(fit) # show results
```{r}
fit <- lm(PIQ ~ Brain + Height + Weight, data=piq)
summary(fit) # show results
```
Equation:
PIQ =  111.35361 + 2.06037 * Brain - 2.73193 * Height + 0.00056 * Weight 

Scatterplot Matrix
```{r}
library(lattice)
pairs(~PIQ+Brain+Height+Weight,data=piq, 
   main=" Scatterplot Matrix")
```
# 3D Scatterplot
```{r}
library(scatterplot3d)
attach(piq)
scatterplot3d(PIQ,Brain,Height,Weight, main="3D Scatterplot")
```
