---
title: "Week 1: Intro to Data"
author: "Anita Owens"
date: "February 9, 2016"
output: html_document
---
UTAustinX: UT.7.10x Foundations of Data Analysis - Part 1
Subsetting: http://www.statmethods.net/management/subset.html
Pre-lab: Cycling Data
```{r}
setwd("/Users/anitaowens/Documents/R")
install.packages("SDSFoundations")
library("SDSFoundations")
bike<-read.csv("BikeData.csv")
```
Prepare for the Analysis
```{r}
# Find the number of students in the dataset
table(bike$student)
#14 students

# Pull out student data into a new dataframe
student <-bike[bike$student==1,]

# Find how often the students ride, using the new dataframe
table(student$cyc_freq)
 #Daily  Less than once a month 
 #                     8                       0 
#Several times per month  Several times per week 
  #                    0                       6 

# Create a vector for the distance variable
distance <-student$distance
#[1]  3.25 10.94  9.34  1.25  9.29  2.77  4.84  6.56  0.85  6.07  3.52
#[12] 13.43  7.19  8.31

# Find average distance ridden
mean(distance)
#6.257857

```
Select the question that will require subsetting the data.
```{r}
riders<-BikeData[ which(BikeData$cyc_freq=="Daily"),]
nrow(riders)
```

What is the average age of daily riders? 
```{r}
mean(riders$age)
round(mean(riders$age), digits = 1)
```
How many of the daily riders are male?
```{r}
sum(with(riders, gender=="M"))
sum(with(riders, gender=="F"))
```
What is the average age of the female daily riders? (Round to 1 decimal place)
```{r}
female<-riders[ which(riders$gender=="F"),]
round(mean(female$age), digits = 1)
```
 What is the average age of the male daily riders? (Round to 1 decimal place.)
 
```{r}
male<-riders[ which(riders$gender=="M"),]
round(mean(male$age), digits = 1)
```
 thirty <- bike[bike$age >= 30,]
 How many daily male riders are age 30 or older?
 ```{r}
thirty<-male[male$age >= 30,]
nrow(thirty)
```

comprehension check
```{r}
workers<-c(10, 2, 6, 12,   14,   15,   15,   24,   15,   25,   3,   12)

mean(workers)
round(mean(workers), digits = 1)
round(sd(workers),digits=2)
round(sd(workers),digits=1)
median(workers)
summary(workers)

range(workers, na.rm = FALSE)
IQR(workers)

plot(workers)
library(ggplot2)
overtime<-as.data.frame(workers)
ggplot(overtime, aes(x=workers)) + geom_histogram(binwidth = .5, color="black", fill="white") + geom_vline(aes(xintercept=mean(overtime, na.rm=T)), color="red", linetype="dashed", size=1)
```


```

