title: 'Week2: Univariate Descriptive Stats Animal Data'
author: "Anita Owens"
date: "March 10, 2016"
output: html_document
---

is this normal
```{r}
AnimalData<-read.csv("AnimalData.csv")
```
animal data
```{r}
#number of variables
ncol(AnimalData)
```
how many were adopted
```{r}
adopt<-AnimalData[1:10,]
sum(with(adopt, Outcome.Type=="Adoption"))
```
plot
```{r}
plot(AnimalData$Animal.Type, AnimalData$Weight, main="Test123")

```
AnimalData Exercises
```{r}
#Find the number of animals that were adopted
table(AnimalData$Outcome.Type)

#Pull out only adopted animals
adopted <- AnimalData[AnimalData$Outcome.Type=="Adoption",]

#Pull out just the days in shelter for the adopted animals
daystoadopt <- adopted$Days.Shelter

#Visualize and describe this variable
hist(daystoadopt)
fivenum(daystoadopt)
mean(daystoadopt)
sd(daystoadopt)
which(AnimalData$Days.Shelter==max(daystoadopt))

#. How would you describe the shape of the distribution of daystoadopt?
#positively skewed-the right tail is longer (right skewed)

IQR(daystoadopt)

mean(adopted$Days.Shelter)
#What was the z-score for this particular animal? Round to the nearest TWO decimal places.
211-mean(adopted$Days.Shelter)
round(181.7353/sd(adopted$Days.Shelter), digits = 2)
#The middle 50% of the distribution were adopted between 8 and 38 days
```
Univariate Descriptive Statistics
```{r}
#subsetting dogs
dogs<-AnimalData[ which(AnimalData$Animal.Type=='Dog' & AnimalData$Age.Intake >= 1),]
nrow(dogs)

#subsetting dats
cats<-AnimalData[ which(AnimalData$Animal.Type=='Cat' & AnimalData$Age.Intake >= 1),]
nrow(cats)
```
weight
```{r}
dogweight<-dogs$Weight
hist(dogweight)

catweight<-cats$Weight
hist(catweight)

summary(dogweight)
summary(catweight)

round(mean(catweight), digits = 1)
round(sd(catweight), digits=2)
```
What is the z-score of a 13 pound adult cat? Round to one decimal point.
```{r}
13-mean(catweight)
round(4.396429/sd(catweight), digits = 1)
```
What proportion of adult cats weigh more than 13 pounds, according to your data? Use the following code to answer this question: 1-pnorm(zcat). Replace "zcat" with your z-score for the cat. Round to three decimal places.
```{r}
zcat<-round(4.396429/sd(catweight), digits = 1)
1-pnorm(zcat)
#[1] 0.01072411
round(1-pnorm(zcat), digits = 3)
```
About 95% of adult cats at the shelter weigh between?
4.8 and 12.4lbs

 ```{r}
quantile(catweight, c(.05, .25, .5, .90, .95))
quantile(catweight, c(.00, .05, .25, .5, .90, .95, 1.00))
```
What was the most common way that dogs arrived in the shelter? (as defined by the “Intake.Type” variable)
```{r}
table(dogs$Intake.Type)
```
What proportion of dogs were brought to the shelter as an owner surrender? (Round to 3 decimal places.)
```{r}
round(58/226, digits = 3)
#answer 0.278

sdogs<-round(zscore/sd(dogs$Intake.Type), digits = 3)
1-pnorm(sdogs)

round(1-pnorm(sdogs), digits = 3)

```
Of the dogs that were brought to the shelter as an owner surrender, how many were returned to their owner?
```{r}
returndogs<-dogs[ which(dogs$Intake.Type=='Owner Surrender' & dogs$Outcome.Type == "Return to Owner"),]
nrow(returndogs)
```
Question3
```{r}
x<-c(2,2,3,5,6,7,9,14,15,18)
y<-c(2,2,3,5,6,8,9,14,15,18)
```