---
title: 'Week3: Bivarate Distributions Bull Riders'
author: "Anita Owens"
date: "March 10, 2016"
output: html_document
---

Week 3 Bivariate Data

Primary Research Question

For the 2013 season, Is there a linear relationship between how often a rider places in the Top 10 and the number of times he stayed on his bull for a full 8 seconds?
```{r}
#import data
BullRiders<-read.csv("BullRiders.csv")
```
Info about the dataset
```{r}
nrow(BullRiders)
```
How many of the first 10 riders in the dataset have been pro for 10 years or more?
```{r}
head(BullRiders)
pro<-BullRiders[1:10,]
plus10<-pro[ which(pro$YearsPro >= 10),]
nrow(plus10)
```
Of the top 15 riders so far in 2015, how many rides were completed by the rider with the fewest buck-outs in 2014?
```{r}
names(BullRiders)
top15<-BullRiders[1:15,]
buck2014<-subset(top15, select=c(Rider, Rank15, BuckOuts14, Rides14))
#sort from lowest to highest
```
Breakdown of Analysis
```{r}
#import data
bull<-read.csv("BullRiders.csv")

#Subset for riders that participated in at least one event in 2013
new_bull <- bull[bull$Events13  > 0 ,]

# Visualize and describe the first variable of interest
hist(new_bull$Rides13)
fivenum(new_bull$Rides13)
mean(new_bull$Rides13)
sd(new_bull$Rides13)

# Visualize and describe the second variable of interest
hist(new_bull$Top10_13)
fivenum(new_bull$Top10_13)
mean(new_bull$Top10_13)
sd(new_bull$Top10_13)

# Create a scatterplot
plot(new_bull$Rides13,new_bull$Top10_13)

# Add line of best fit
abline(lm(new_bull$Top10_13~new_bull$Rides13))

# Calculate the correlation coefficient
cor(new_bull$Rides13,new_bull$Top10_13)

# Create a correlation matrix 
vars <- c("Top10_13", "Rides13")
cor(new_bull[,vars])
round(cor(new_bull[,vars]), digits = 3)
```
Which cases will be selected by this line of code?
```{r}
new_bull <- bull[bull$Events13 > 0 ,]
#riders that have completed more than zero Events in 2013
```
In the above scatterplot, why does the line of best fit seem to not be going through the center of the scatterpot? (Refer to the code below and the dataset in R for help.)
```{r}
plot(bull$Events12, bull$BuckOuts12)
abline(lm(bull$Events12~bull$BuckOuts12))

#corrected code
plot(bull$Events12, bull$BuckOuts12)
abline(lm(bull$BuckOuts12~bull$Events12))
```
A correlation matrix allows you to calculate multiple correlation coefficients at a time. Here, we are only asking for the correlation between Rides13 and Top10_13. If you wanted to include other variables as well, how would you do that?
```{r}
# Create a correlation matrix
vars <-c("Top10_13", "Rides13")
cor(new_bull[,vars])

```
On average, a bull rider in 2013 has how many rides? (Report the median becuase the histogram is not symmetrical.)
```{r}
median(new_bull$Rides13)
```
1b. These bull riders made it into the Top 10 an average of ________ times in 2013. (Hint: Remember again that the histogram is not symmetrical.)
```{r}
mean(new_bull$Top10_13)
```
Use this code to help identify this rider:
#identify a specific record 
```{r}
#identify a specific record
which(new_bull$Top10_13==2 & new_bull$Rides13==22)
record<-new_bull[4,]
```
Create a dataset which contains riders that participated in at least one event in 2012. Call the dataset new_bull12.
```{r}
new_bull12<-BullRiders[ which(BullRiders$Events12 > 0),]
```
Make a histogram to visualize the distribution of Earnings for 2012.
```{r}
hist(new_bull12$Earnings12)
```
Generate the appropriate descriptive statistics for this distribution
```{r}
fivenum(new_bull12$Earnings12)
mean(new_bull12$Earnings12)
sd(new_bull12$Earnings12)
median(new_bull12$Earnings12)
```
1b. What was the average amount earned by a bull rider? (Choose the appropriate measure of center; report without a $ sign and round to the nearest whole number.)
```{r}
round(median(new_bull12$Earnings12), digits = 0)
```
What was the highest amount earned by a bull rider? (Report without a $ sign and round to the nearest whole number.)
```{r}
which.max(new_bull12$Earnings12)

#see which row is the highest earner
high<-new_bull12[4,]
#Silvano Alves

#subsetting income earnings for 2012
earnings<-subset(new_bull12, select = c(Rider, Rank15, Earnings12))
#then sort from highest to lowest

round(1464475.61, digits = 0)
```
Make a Scatterplot of Earnings and Ride Percentage
```{r}
plot(new_bull12$Earnings12,new_bull12$RidePer12)
```
correlation matrix
```{r}
# Create a correlation matrix
vars <-c("Earnings12", "RidePer12")
round(cor(new_bull12[,vars]), digits = 3)
```
Create a Scatterplot of Earnings and Cup Points
```{r}
plot(new_bull12$Earnings12,new_bull12$CupPoints12)
abline(lm(new_bull12$CupPoints12~new_bull12$Earnings12), col="red")
```
correlation matrix
```{r}
vars <-c("Earnings12", "CupPoints12")
round(cor(new_bull12[,vars]), digits = 3)
```
Outliers and Influential Points

An outlier can have a significant impact on the correlation coefficient. Sometimes it is important to remove these points to examine the size of this impact. Run this code to identify the extreme data value in Earnings:
```{r}
# identify specific case
which(new_bull12$Earnings12 == max(new_bull12$Earnings12))
```
plot
```{r}
plot(new_bull12$RidePer12, new_bull12$Earnings12)
abline(lm(new_bull12$Earnings12~new_bull12$RidePer12), col="red")
```
Let's remove this data point from the dataset to assess what kind of impact, if any, it had on our correlation analysis. Run this code:
```{r}
#Subset the data
nooutlier <- new_bull12[new_bull12$Earnings12 < 1000000 ,] 
```
Then rerun the correlation matrix and the scatterplots to see the difference. Make sure to use the new dataframe (nooutlier) that you just created.
```{r}
#correlation matrix
vars <-c("Earnings12", "RidePer12")
round(cor(nooutlier[,vars]), digits = 3)

#scatterplot
plot(nooutlier$RidePer12, nooutlier$Earnings12)
abline(lm(nooutlier$Earnings12~nooutlier$RidePer12), col="red")
```
After removing the outlier, what was the new correlation of Earnings and Cup Points for 2012? (Round to three decimals)
```{r}
#correlation matrix
vars <-c("Earnings12", "CupPoints12")
round(cor(nooutlier[,vars]), digits = 3)
```
Answer the question and support your answer with statistics:

An initial examination of the relationships between Ride Percentage (RidePer) and Earnings, and Cup Points (CupPoints) and Earnings showed that Cup Points had the  
stronger - This answer is correct. stronger relationship to Earnings. Ride Percentage and Earnings showed a correlation value of   
0.593 - This answer is correct. 0.593 while Cup Points and Earnings had a correlation value of   
0.657 - This answer is correct. 0.657 . Visual examination showed an   
outlier - This answer is correct. outlier in both relationships—a rider who has earned over 1 million dollars. Removal of this increased the initial relationship: Ride Percentage and Earnings now had a correlation value of  
0.804 - This answer is correct. 0.804 and Cup Points and Earnings had a correlation value of  
0.893 - This answer is correct. 0.893 .   
Cup Points - This answer is correct. Cup Points still had the higher relationship to Earnings. Visual examination showed a  
good linear relationship - This answer is correct. good linear relationship for both Ride Percentage and Cup Points, indicating the correct use of the correlation coefficient.

PROBLEM SET
Subset the dataset for riders that had at least 1 ride in the 2014 season. Call this dataset new_bull.
```{r}
new_bull<-BullRiders[ which(BullRiders$Rides14 >= 1),]
```
Create a new variable
```{r}
RidesPerEvent14 <- new_bull$Rides14/new_bull$Events14
```
Make a histogram of your "rides per event" variable and find the five-number summary for your "rides per event" variable.
```{r}
hist(RidesPerEvent14)
round(fivenum(RidesPerEvent14), digits = 2)
```
Create a scatterplot of "rides per event" and yearly ranking (defined by the "Rank14" variable) and add a line of best fit. Which of the following best describes the relationship between these two variables?
```{r}
plot(RidesPerEvent14, new_bull$Rank14)
abline(lm(new_bull$Rank14~RidesPerEvent14), col="red")
```
What is the correlation coefficient for rides per event and yearly ranking? (Report to 3 decimal places)
```{r}
#adding the RidersPerEvent14 column to dataframe
test<-cbind(new_bull, RidesPerEvent14)

#correlation matrix
vars <-c("RidesPerEvent14", "Rank14")
round(cor(test[,vars]), digits = 3)
```
Suppose that college GPA and graduate school GPA have a correlation coefficient of 0.75. Based on this, what proportion of variation in graduate school GPA is left unexplained after taking college GPA into account? (Report to 4 decimal places)?

If |r| is between 0.5 and 0.85, there is a moderate correlation
```{r}
.75*.75
1-.5625
#56% of graduate school GPA can be attributed to college GPA
#Answer: .4375
```
Using the dataset below, find the correlation coefficient between time spent studying and exam grade.
```{r}
#import data
study <- read.csv("~/Documents/R/study.csv")

#correlation matrix
vars <-c("Minutes", "Grade")
round(cor(study[,vars]), digits = 3)
```
Approximately what percentage of the variation in exam scores can be explained by the amount of time that each student studied? (Round to whole number without a % sign.)
```{r}
.597*.597
0.356409
round(0.356409*100, digits = 0)
```
Create a scatterplot of the data (exam grades and time spent studying). What is the value of the outlier (the student that got a high grade but didn't study very long)?
```{r}
#plot(study$Grade, study$Minutes)
plot(study$Minutes, study$Grade)
```
When the outlier is removed, what is the new value of r? (Round to 3 decimal places.)
```{r}
#remove outlier
nooutlier<-study[ which(study$Grade !=92),]
nrow(nooutlier)

#correlation matrix
vars <-c("Minutes", "Grade")
round(cor(nooutlier[,vars]), digits = 3)
```
Z-score: If an adult has a z-score of -1.5, how many hours of sleep does this person get per night? (Report to 2 decimal places.)
```{r}

round(4.59-6.7/1.1, digits = 2)

#solved for x by hand
#hours of sleep=5.05 - 6.7 / 1.1 = -1.5 z-score

```
What proportion of adults sleep longer than 4.5 hours per night? (Report to 3 decimal places.)
answer: 0.977
```{r}
1-pnorm(zscore)
round(1-pnorm(zscore), digits = 3)

#z-score
zscore<- (4.5 - 6.7)/1.1
round(1-pnorm(zscore), digits = 3)
#answer:  0.977
```
What proportion of adults sleep between 5.38 and 8.79 hours of sleep? (Report to 3 decimal places.)
answer:.856
```{r}
zscore2<-(5.38-6.7) / 1.1
round(1-pnorm(zscore2), digits = 3)
#.885

#What is the probability that a value with a z-score of -0.7109091 will occur in a normal distribution?

#solution:
#Scroll up to the table above and find “ -0.7” on the left or right side. Now move across the table to “0.01” on the top or bottom, and record the value in the cell: 0.2389. That tells us that 23.89% of values in the set are at or below a z -score of -0.7109091. 


zscore3<-(8.79-6.7)/ 1.1
round(1-pnorm(zscore3), digits = 3)
#0.029
#What is the probability that a value with a z-score of 2.699091 will occur in a normal distribution?

#solution:
#Scroll up to the table above and find “ 2.6” on the left or right side. Now move across the table to “0.09” on the top or bottom, and record the value in the cell: 0.9964. That tells us that 99.64% of values in the set are at or below a z -score of 2.699091.

answer<-0.885-0.029
```

Probability Example

Example B
What is the probability that a value with a z-score greater than 1.53 will occur in a normal distribution? 
Solution
Scroll up to the table of z-score probabilities again and find the intersection between 1.5 on the left or right and 3 on the top or bottom, record the value in the cell: 0.937 .

That decimal lets us know that 93.7% of values in the set are below the z-score of 1.53. 

To find the percentage that is above that value, we subtract 0.937 from 1.0 (or 93.7% from 100%), to get 0.063 or 6.3%.
```{r}

```
What is the z-score of the price of a pair of skis that cost $247, if the mean ski price is $279, with a standard deviation of $16?
```{r}
zscoreski<-(247-279)/16
#zscore=-2
```
