---
title: "xlsx package"
author: "Anita Owens"
date: "February 24, 2016"
output: html_document
---
https://www.datacamp.com/community/tutorials/r-tutorial-read-excel-into-r

load package
```{r}
library(xlsx)

```
example of reading xlsx sheets

file : the path to the file to read
sheetIndex : a number indicating the index of the sheet to read; e.g : use sheetIndex=1 to read the first sheet
header : a logical value. If TRUE, the first row is used as the names of the variables
colClasses : a character vector that represents the class of each column
```{r}
 file <- system.file("tests", "test_import.xlsx", package = "xlsx")
    res <- read.xlsx(file, 2)  # read the second sheet
```
example of writing xlsx sheets

write.xlsx(x, file, sheetName="Sheet1", 
  col.names=TRUE, row.names=TRUE, append=FALSE)

write.xlsx2(x, file, sheetName="Sheet1",
  col.names=TRUE, row.names=TRUE, append=FALSE)
```{r}
file <- paste(tempfile(), "xlsx", sep=".")
    write.xlsx(USArrests, file=file)
```
sex survey data
```{r}
read.xlsx("sex_survey.xlsx", sheetIndex=1, header=TRUE, colClasses=NA)

#creating a data frame #this took hours to do
sex<-read.xlsx("sex_survey.xlsx", sheetIndex=1, header=TRUE, colClasses=NA)

read.xlsx2(file, sheetIndex, header=TRUE, colClasses="character")

#save a data object
save(sex, file = "saveddf.RData")

#write file
write.csv(sex, file = "savedsexfile.csv")
#clean up mess
unlink("savedsexfile.csv")
```
sd data
```{r}
#reading xlsx file into a dataframe
df<-read.xlsx("sd.xlsx", sheetIndex=1, header=TRUE, colClasses=NA)

#2nd option
read.xlsx2(file, sheetIndex, header=TRUE, colClasses="character")
```
create a new workbook
```{r}
# create a new workbook for outputs
# possible values for type are : "xls" and "xlsx"
wb<-createWorkbook(type="xlsx")
```
https://flowingdata.com/2012/12/17/getting-started-with-charts-in-r/
