---
title: "RFM Sonetel: Statistical Segmentation"
author: "Anita Owens"
date: "September 2, 2016"
output: html_document
---
read csv file
```{r}
sont<-read.csv("SontData.csv")
str(sont)
```
# Add headers and interpret the last column as a date, extract year of purchase

```{r}
sont$last_payment_date = as.Date(sont$last_payment_date, "%Y-%m-%d")
sont$days_since= as.numeric(difftime(time1 = "2016-09-01",
                time2 = sont$last_payment_date, units                         = "days"))
str(sont)
```
# Compute key marketing indicators using SQL language
```{r}
library(sqldf)
```
# Compute recency, frequency, and average purchase amount
```{r}
customers = sqldf("SELECT accountID,
                          MIN(days_since) AS 'recency',
                          num_of_pymt AS 'frequency',
                          avg_pymt_amount AS 'amount'
                   FROM sont GROUP BY 1")
```
Statistical Segmentation
Explore Data
```{r}
# Explore the data
head(customers)
summary(customers)
hist(customers$recency)
hist(customers$frequency)
hist(customers$amount)
hist(customers$amount, breaks = 100)
hist(customers$amount, breaks = 200)
```
# --- PREPARING AND TRANSFORMING DATA ----------------------
# Copy customer data into new data frame/carbon copy
```{r}
new_data = customers
```
# Remove customer id as a variable, store it as row names
```{r}
head(new_data)
row.names(new_data) = new_data$accountID
new_data$accountID = NULL
head(new_data)
```
# Take the log-transform of the amount, and plot
```{r}
new_data$amount = log(new_data$amount)
hist(new_data$amount)
```
# Standardize variables
```{r}
new_data = scale(new_data)
head(new_data)
```
# Compute distance metrics on standardized data
# This will likely generate an error on most machines
```{r}
# d = dist(new_data)
```
no sampling needed
# Take a 10% sample
```{r}
#sample = seq(1, 18417, by = 10)
#head(sample)
#customers_sample = customers[sample, ]
#new_data_sample  = new_data[sample, ]
```
# Compute distance metrics on standardized data
```{r}
d = dist(new_data)
```
# Perform hierarchical clustering on distance metrics
```{r}
c = hclust(d, method="ward.D2")
```
# Plot de dendogram
```{r}
plot(c)
```
# Cut at 9 segments
```{r}
members = cutree(c, k = 9)
```
# Show 30 first customers, frequency table
```{r}
members[1:30]
table(members)
```
# Show profile of each segment
#need to fix
```{r}
#aggregate(new_data[, 2:4], by = list(members), mean)

aggregate(new_data[, 2:4], by=list(members), mean)
```
MANAGERIAL SEGMENTATION------>
# Add headers and interpret the last column as a date, extract year of purchase
```{r}
sont<-read.csv("SontData.csv")
str(sont)
sont$last_payment_date = as.Date(sont$last_payment_date, "%Y-%m-%d")
sont$year_of_purchase = as.numeric(format(sont$last_payment_date, "%Y"))
sont$days_since= as.numeric(difftime(time1 = "2016-09-01",
                time2 = sont$last_payment_date, units = "days"))

```
display data after transformation
```{r}
head(sont)
summary(sont)
```
# Compute key marketing indicators using SQL language
```{r}
library(sqldf)

# Compute recency, frequency, and average purchase amount
customers_2015 = sqldf("SELECT accountID,
                               MIN(days_since) AS 'recency',
                               MAX(days_since) AS 'first_purchase',
                               num_of_pymt AS 'frequency',
                               avg_pymt_amount AS 'amount'
                        FROM sont GROUP BY 1")
head(customers_2015)
```
#explore the data
```{r}
head(customers_2015)
summary(customers_2015)
hist(customers_2015$recency)
hist(customers_2015$frequency)
hist(customers_2015$amount)
hist(customers_2015$amount, breaks = 100)
```



